---
title: 'Statistical report: Resistant hypertension in SwedeHF'
author: 'Statistician: Lina Benson'
  
date: "`r Sys.Date()`"
output:
  pdf_document:
    fig_caption: yes
    fig_height: 7
    fig_width: 7
    number_sections: yes
link-citations: yes
bibliography: references.bib
nocite: '@*'
---

\newpage 
\tableofcontents 
\listoffigures
\listoftables
\newpage


```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE, include = TRUE, comment = "",
  warning = FALSE, message = FALSE, fig.pos = "H",
  fig.path = "../output/figs/"
)
options(knitr.kable.NA = "")
```

```{r adjust_directory_if_needed, include=FALSE}
# Uncomment lines below if rmd file is placed in a subdirectory
knitr::opts_knit$set(root.dir = normalizePath("../"))
```

```{r load_project}
# 1. Set options in config/global.dcf
# 2. Load packages listed in config/global.dcf
# 3. Import functions and code in lib directory

ProjectTemplate::reload.project()

cacheon <- TRUE
```             

# Data handling

## Data source

SHFDB3, https://kiheartfailure.github.io/shfdb3/, v 3.2.3. 

## Inclusion/exclusion criteria

```{r flow}
mykable(flow, caption = "Flowchart")
```

First patient in: `r min(pdata$shf_indexdtm)` and last patient in: `r max(pdata$shf_indexdtm)`.  

The median age (IQR) is `r pdata %>% summarise(med = dF(median(shf_age), dig = 1),
                                             q1 = dF(quantile(shf_age, probs = 0.25), dig = 1),
                                             q3 = dF(quantile(shf_age, probs = 0.75), dig = 1)) %>%
                                   mutate(out = paste0(med, " (", q1, "-", q3, ")")) %>%
                                   pull(out)` and 
`r pdata %>% count(shf_sex) %>%
  mutate(perc = dF(n / sum(n) * 100, 1)) %>%
  filter(shf_sex == "Female") %>%
  pull(perc)`% females.    
  
## Created variables 

```{r atc}
mykable(metalm, caption = "Treatments from DDR")
```

To extract hypertension from the NPR ICD-10 I10-I15 were used. The maximum possible 
duration was set to 9 years (in order to give all patients possibility to 
have the same amount of duration, irrespectively of when they were in included 
in SwedeHF, since ICD-10 was introduced in 97/98 in the NPR). 

The definition of HT is: 

- if shf_bpsys >= 140 OR (shf_sos_com_diabetes == "Yes" AND shf_bpsys >= 135) 
AND ddr_ccbl == "Yes" AND ddr_diuretic == "Yes" AND ddr_rasarni == "Yes" -> Resistant HT

- if shf_bpsys >= 140 OR shf_sos_com_diabetes == "Yes" & shf_bpsys >= 135 without meds as above -> Non-resistant HT

- else Normal

# Statistical analysis 

## General

All analyses were performed using `r sessionInfo()$R.version$version.string` [@r]. 
The level of significance is set to 5%, two-sided. No adjustment for multiple 
comparisons were made and therefore the results should be viewed with some care.

## Missing data

Missing data was imputed with multiple imputation (n = 10) using mice [@mice]. 
Variables included in the model are indicated in 
Table \ref{tab:tab1}. The primary outcome, first heart failure hospitalization/cv death,
was included as the Nelson-Aalen estimator. 

## Baseline characteristics

```{r, child = "../src/tab1.Rmd"}

```

```{r, child = "../src/barplot.Rmd"}

```

\clearpage
\newpage 

## Associations with HT

Logistic regression models were used to evaluate the association 
between baseline characteristics and HT in each EF group separately using the interaction term for the respective variable and EF. 
Different models were used for Non-resistant HT vs Normal and Resistant HT vs Normal. 

```{r, child = "../src/ortab.Rmd"}

```

\clearpage
\newpage

## Association HT vs outcomes

The following outcomes are considered: 

- Cardiovascular mortality/first HF hospitalization (primary endpoint)
- First HF hospitalization
- Cardiovascular mortality
- All-cause mortality

Incidence per 1000 py was calculated with 95% Poisson confidence intervals. 

The outcomes are presented with the 1 - Kaplan-Meier curves (for outcomes including all-cause mortality) 
or cumulative incidence curves treating death as a competing event and tested with the log-rank test (note that the log-rank test is theoretically 
redundant since the cox regression gives the same information). 

Cox proportional hazards regressions were used to evaluate the association 
between HT and the respective outcomes. The interaction between EF and HT was included enabling presentation in each 
EF group separately and at the same time being ale to statistically test for differences between EF groups.   

Data were censored at 2019-12-31, after 3 years of follow-up 
or death/emigration, so this assumes that the risk of cv mortality and hf hospitalization 
would have been the same for a censored patient had the censoring not occurred. 

The median (min-max) follow-up is 
`r pdata %>% summarise(med = dF(median(sos_outtime_death / 365.25 * 12), dig = 1),
                                             min = dF(min(sos_outtime_death / 365.25 * 12), dig = 1),
                                             max = dF(max(sos_outtime_death / 365.25 * 12), dig = 1)) %>%
                                   mutate(out = paste0(med, " (", min, "-", max, ")")) %>%
                                   pull(out)` months for a total of 
                                   `r pdata %>% summarise(sumpy = dF(sum(sos_outtime_death) / 365.25, dig = 0)) %>%
                                   pull(sumpy)` patient-years of follow-up.

### Assumptions

The proportional hazards assumption was investigated using the scaled Schoenfeld 
residuals (cox.zph in [@survival-package]) for the primary outcome. 
There were no signs of non-proportional hazards. 

\clearpage
\newpage

The median (min-max) follow-up is 
`r pdata %>% summarise(med = dF(median(sos_outtime_death / 365.25 * 12), dig = 1),
                                             min = dF(min(sos_outtime_death / 365.25 * 12), dig = 1),
                                             max = dF(max(sos_outtime_death / 365.25 * 12), dig = 1)) %>%
                                   mutate(out = paste0(med, " (", min, "-", max, ")")) %>%
                                   pull(out)` months for a total of 
                                   `r pdata %>% summarise(sumpy = dF(sum(sos_outtime_death) / 365.25, dig = 0)) %>%
                                   pull(sumpy)` patient-years of follow-up.

```{r, child = "../src/km.Rmd"}

```

```{r, child = "../src/coxtab.Rmd"}

```

\clearpage
\newpage

# Reproducibility

## R session information {#sessioninfo}

```{r sessinfo}
sessionInfo()
```

## R code

The R code for all data handling and statistical analyses are found: 
https://github.com/KIHeartFailure/resistantht. On publication
the repository will be made public so as to 
link to it from the resulting article for increased transparency and code sharing.
No data or output is stored in the repository. 

# References
